### 2013-04-02 doc/final_report/design.tex
Made some changes to design chapter as suggested by Amanda and Maria

Given the project objectives identified above, four primary use cases were
identified for Partridge as shown in Figure \ref{fig:use_cases}. Users can be
identified as researchers and people interested in finding and reading
scientific papers that interested them. This would involve querying the
database for relevant papers, viewing metrics on the papers that have been used
to classify them and also downloading the original paper for reading. A subset
of users were identified as authors who might be interested in adding their own
papers to the Partridge instance.

@@

One of the main requirements for a web server is that it responds to requests
relatively quickly, ideally within 10 seconds or less. If they don't, the user
may get impatient and give up trying to use the system, or in some cases, their
web browser may timeout and stop trying to load the page. Many of the processes
involved in extracting meaningful CoreSC information from papers are very slow
slow. This means that processing papers during web requests is not really
feasible, since it may take several minutes for a single paper to be annotated
and classified. Therefore, a third core component was identified: a paper
preprocessor service that runs on the server and converts, annotates and
classifies papers as soon as they are uploaded.

@@

To further increase development speed, a Python framework for developing web
applications called \emph{Flask\cite{flask2012}} would be used to build the web
backend.

@@

Due to the large number of popular web browsers on the market, implementing
JavaScript in a slightly different way, Javascript can be very difficult to
write in a cross-browser compatible way.  To maximise compatibility and reduce
development time, it was decided that libraries such as jQuery which are
designed to carry out HTML manipulation behaviours in a uniform way cross all
compatible browsers.  As of February 2013, it is estimated that Chrome and
Firefox are currently the most popular web browsers, holding 50\% and 29.6\% of
the market share respectively\cite{browserstats2013}. Internet Explorer is the
third most popular browser. However, there are several well known compatibility
issues with Internet Explorer, Microsoft themselves condemning version 6.0 of
their browser\cite{ie6death}. For this reason, it was decided that Partridge
would only support Firefox and Chrome. A working interface in modern editions
of the Internet Explorer family of browsers would be a bonus.  However, no time
was allocated to ensuring compatibility of Partridge with Internet Explorer.

@@

\begin{figure}[!ht]
\center
\includegraphics[width=0.9\textwidth]{images/design/components_high_level.png}
\caption{High Level Component Layout for Partridge}
\label{fig:high_level_components}
\end{figure}

@@

\subsubsection{PDF Conversion}

@@

Once the paper has been split, SAPIENTA is used to annotate each sentence with
a relevant CoreSC allocation. These labels are calculated based upon the
sentence's location within the paper as well pairs and triplets of words found
consecutively within the sentence. After this stage, the paper is passed
forwards to be further classified by Partridge.

@@

The final stage in the preprocessor pipeline is paper classification. During
this stage, features are extracted from the papers and used to classify the
paper in three different ways. The paper's type: case study or review paper
etc. is determined using the presence or absence of different CoreSC labels
within the paper and their relative proportions. The type of paper is
determined by the presence of domain specific nouns and adjectives within the
paper. For example, a biology paper might contain words such as enzyme or amino
acid. The result of the paper, i.e. whether the paper was conclusive and if the
results were positive or negative, are detected through the presence or absense
of words with a strong sentiment polarity as discussed by Wilson et al
(2005)\cite{Wilson05Polarity}. The results from the whole process are then
passed into the data store and saved.

@@

Each of the independent processes described above generate new files which are
temporarily stored on the server until all of the sub-processes are complete.
Once a paper analysis cycle has finished, the preprocessor triggers a cleanup
operation that removes intermediate files from the system, storing only the
initially uploaded file (which may be a PDF) and the final annotated file in a
permanent location on the filesystem.  The whole analysis process for a single
paper is also wrapped within an error handling block that automatically run the
cleanup routine if something goes wrong. This prevents the system filling up
with invalid files. This mechanism is shown in Figure
\ref{fig:preprocessor_overview}.

@@

\subsection{ Data Storage }
\label{sec:db_layout}
It is expected that a given Partridge instance may grow in size to contain
thousands of individual papers that have been annotated and are available to
the web backend for inclusion into a user's search results. Reading in and
parsing each individual XML file stored in the system is not really feasible
when handling a HTTP request from the web frontend, this sort of process would
take a long time to execute. To provide fast search and retrieval of paper
data, a Relational Database Management server (RDBMS) was chosen for storage of
papers that have been preprocessed using the system described above. An
overview of the database schema is illustrated in Figure \ref{fig:e-r-diagram}.

@@

As discussed above, two or more files per paper are always stored within
partridge so that a user can retrieve them for reading. Therefore files are
kept in their own table along with a record of their type (eg. XML or PDF).

@@

The Flask web framework uses a model view controller-like architecture. Each
behaviour of the server: handling paper uploads, displaying paper query
results, allowing users to download papers, can be considered a `view'. Views
register themselves with Flask under a specific URL. When flask receives a
request to any such URL, it automatically routes the request through to the
View object\cite{flask2012}.

@@

The key responsibility of the query view is to translate incoming HTTP requests
into search queries, retrieve a set of papers from the RDBMS  and return human
readable results to the user. One of the most important steps in this process
is to cleanse the incoming data and prevent SQL injection attacks that
unscrupulous users may attempt to use to damage or purge the Partridge
database. The system must also be capable of interpreting data retrieved from
the database into HTML to be returned to the user's browser. The server
mechanism utilises Flask's built in templating mechanism in order to quickly
transform Python-specific data structures into valid HTML markup and return it
to the user\cite{flask2012}.

@@

The search or `Keywords' section of the form shows a set of controls for
specifying terms that must appear within sections of a document for it to show
up in the search results. The user enters a search term in the input box on the
top right of the form and then selects which section of the paper the term
should appear within. This could be any of the CoreSC zones within the paper or
within the abstract. The user could also specify a specific author's name. Once
they are happy with their keywords and section, they click ``Add" and the
keyword/section pair are inserted into the table on the bottom right. If they
wish to remove a query constraint from this table, they can select it and click
the ``Delete" button.

@@

The paper's title and author are displayed prominently in the top left hand
corner of the page. The full abstract text is shown so that the user can read
it and determine the paper's relevance to them. There is also a graph showing
each of the CoreSC types and proportion within the paper. If the user is
interested in the paper, they can click the Download link at the top of the
page to access the annotated XML file stored in the system.

@@

Partridge has been developed using an iterative methodology and therefore some
of the designs descussed in this chapter have been modified during the
development of the project. However, these designs did provide an excellent
starting point for the implementation of the project.

>>>>

Given the project objectives identified in \ref{sec:objectives}, four primary
use cases were identified for Partridge as shown in Figure \ref{fig:use_cases}.
Users of Partridge are likely to be researchers who want to find and read
scientific papers. This would involve querying the database for relevant
papers, viewing metrics on the papers that have been used to classify them and
also downloading the original paper for reading.

Authors are the subset of users who may remain interested in using the query
functionality of Partridge, but also may wish to upload papers to Partridge and
add them to the corpus.

@@

One of the main requirements for a web server is that it responds to requests
relatively quickly, ideally within 10 seconds or less. If the response is any
slower than this, the user may get impatient and give up trying to use the
system, or in some cases, their web browser may timeout and stop trying to load
the page. Many of the processes involved in extracting meaningful CoreSC
information from papers are very slow. This means that processing papers
during web requests is not really feasible, since it may take several minutes
for a single paper to be annotated and classified. Therefore, a third core
component was identified: a paper preprocessor service that runs on the server
and converts, annotates and classifies papers as soon as they are uploaded.

@@

To further increase development speed, a Python framework for developing web
applications called \emph{Flask} would be used to build the web backend. Flask
is a ``...micro webdevelopment framework for Python.\cite{flask2012}" It
contains a set of utilities and tools for building web applications quickly and
effectively in Python, taking care of low level behaviours such as TCP socket
communication and HTTP requests. These capabilities are presented in a
Model-View-Controller pattern. Flask's capabilities and use within Partridge is
discussed further in Section \ref{sec:backend}.

@@

Due to the large number of popular web browsers on the market, implementing
JavaScript in a slightly different way, Javascript can be very difficult to
write in a cross-browser compatible way. To maximise compatibility and reduce
development time, it was decided jQuery would be used in developing the
JavaScript frontend behaviours. jQuery is a library of utility functions
designed for manipulating HTML documents in a uniform way across all compatible
browsers \cite{jquery2013}. As of February 2013, it is estimated that Chrome and
Firefox are currently the most popular web browsers, holding 50\% and 29.6\% of
the market share respectively\cite{browserstats2013}. Internet Explorer is the
third most popular browser. However, there are several well known compatibility
issues with Internet Explorer, Microsoft themselves condemning version 6.0 of
their browser\cite{ie6death}. For this reason, it was decided that Partridge
would only support Firefox and Chrome. A working interface in modern editions
of the Internet Explorer family of browsers would be a bonus.  However, no time
was allocated to ensuring compatibility of Partridge with Internet Explorer.

@@

\begin{figure}[!ht]
\center
\includegraphics[width=0.6\textwidth]{images/design/components_high_level.png}
\caption{High Level Component Layout for Partridge}
\label{fig:high_level_components}
\end{figure}

@@

\subsubsection{PDF to XML Conversion}

@@

Once the paper has been split, SAPIENTA is used to annotate each sentence with
a relevant CoreSC label. These labels are calculated based upon the
sentence's location within the paper as well pairs and triplets of words found
consecutively within the sentence. After this stage, the paper is passed
forwards to be further classified by Partridge.

@@

The final stage in the preprocessor pipeline is paper classification. During
this stage, features wil be extracted from the papers and used to classify the
paper in three different ways. The paper's type: case study or review paper
etc. will be determined using the presence or absence of different CoreSC labels
within the paper and their relative proportions. The subject of paper will be
determined by the presence of domain specific nouns and adjectives within the
paper. For example, a biology paper might contain words such as enzyme or amino
acid. The result of the paper, i.e. whether the paper was conclusive and if the
results were positive or negative, will be detected through the presence or absense
of words with a strong sentiment polarity as discussed by Wilson et al
(2005)\cite{Wilson05Polarity}. The results from the whole process are then
passed into the data store and saved.

@@

Each of the independent processes described above generate new files which are
temporarily stored on the server until all of the sub-processes are complete.
Once a paper analysis cycle has finished, the preprocessor triggers a cleanup
operation that removes intermediate files from the system, storing only the
initially uploaded file (which may be a PDF) and the final annotated file in a
permanent location on the filesystem.  The whole analysis process for a single
paper is also wrapped within an error handling block that automatically
executes the cleanup routine if something goes wrong. This prevents the system
filling up with invalid files. This mechanism is shown in Figure
\ref{fig:preprocessor_overview}.

@@

\subsection{ Data Storage }
\label{sec:db_layout}
It is expected that a given Partridge instance may grow in size to contain
tens of thousands of individual papers that have been annotated and are available to
the web backend for inclusion into a user's search results. Reading in and
parsing each individual XML file stored in the system is not really feasible
when handling a HTTP request from the web frontend, this sort of process would
take a long time to execute. To provide fast search and retrieval of paper
data, a Relational Database Management server (RDBMS) was chosen for storage of
papers that have been preprocessed using the system described above. An
overview of the database schema is illustrated in Figure \ref{fig:e-r-diagram}.

@@

As discussed above, two or more files per paper are always stored within
Partridge so that a user can retrieve them for reading. Therefore files are
kept in their own table along with a record of their type (eg. XML or PDF).

@@

The Flask web framework uses a model view controller-like architecture. Each
behaviour of the server: handling paper uploads, displaying paper query
results, allowing users to download papers, can be considered a `view'. Views
register themselves with Flask under a specific URL. When Flask receives a
request to any such URL, it automatically routes the request through to the
View object\cite{flask2012}.

@@

The key responsibility of the query view is to translate incoming HTTP requests
into search queries, retrieve a set of papers from the RDBMS  and return human
readable results to the user. One of the most important steps in this process
is to cleanse the incoming data and prevent SQL injection attacks that
unscrupulous users may attempt to use to damage or purge the Partridge
database. These are handled transparently by the SQLAlchemy library as
discussed in Section \ref{sec:dbandsqlalchemy}. The system must also be capable
of interpreting data retrieved from the database into HTML to be returned to
the user's browser. The server mechanism utilises Flask's built in templating
mechanism in order to quickly transform Python-specific data structures into
valid HTML markup and return it to the user\cite{flask2012}.

@@

The search or `Keywords' section of the form shows a set of controls for
specifying terms that must appear within sections of a document for it to show
up in the search results. The user enters a search term in the input box on the
top right of the form and then selects which section of the paper the term
should appear within. This could be any of the CoreSC labels within the paper or
within the abstract. The user could also specify a specific author's name. Once
they are happy with their keywords and section, they click ``Add" and the
keyword/section pair are inserted into the table on the bottom right. If they
wish to remove a query constraint from this table, they can select it and click
the ``Delete" button.

@@

The paper's title and author are displayed prominently in the top left hand
corner of the page. The full abstract text is shown so that the user can read
it and determine the paper's relevance to them. There is also a graph showing
each of the CoreSC types and proportion within the paper (the example chart only
shows three CoreSC label types for brevity). If the user is
interested in the paper, they can click the Download link at the top of the
page to access the annotated XML file stored in the system.

@@

Partridge has been developed using an iterative methodology and therefore some
of the designs discussed in this chapter have been modified during the
development of the project. However, these designs did provide an excellent
starting point for the implementation of the project.
